---
title: "Multiple Pages"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    self_contained: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(readxl)
library(haven) #for loading other datafiles (SAS, STATA, SPSS, etc.)
library(stringr) # if you have not installed this package, please install it.
library(ggplot2) # if you have not installed this package, please install it.
library(sf) 
library(mapview)
#Loading colorblind-friendly color map package: viridisLite
library(viridis)
```

# Page 1

## Column {data-width="650"}

### Starbuck Location Live Map

```{r}
#read the data: 
starbucks <- read_csv("https://raw.githubusercontent.com/libjohn/mapping-with-R/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv", show_col_types = FALSE)
#head(starbucks)

##Create a column of "gross profit" and assign random numbers to it:
starbucks <- starbucks |>
  mutate(gross_profit = sample(1000:10000, size = n(), replace = TRUE))
#head(starbucks)

starbucksMA <- starbucks  |>
  filter(State %in% c("MA"))

starbucks_sf <- starbucksMA |>
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

mapview(starbucks_sf, col.regions = "red", cex = "gross_profit", legend = TRUE, map.types = "OpenStreetMap")
#col.region: Sets all bubbles to a fixed color
#cex:sets the bubble size to be proportional to a specific column


```

## Column {data-width="350"}

### Dataset

Source: [Starbucks Location in the US](https://raw.githubusercontent.com/libjohn/mapping-with-R/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv)

In this visualization, we subset the data by keeping only starbucks in Massachusetts.

```{r}

```

### Findings

Starbucks are more condensed in urban area, such as the greater Boston and Springfield.

```{r}

```

# Page 2 {data-orientation="rows"}

## Row {data-height="600"}

### Chart 1

```{r}
```

## Row {data-height="400"}

### Chart 2

```{r}
```

### Chart 3

```{r}
```
